---
title: "Titanic Survival Predictor"
subtitle: "Developing Data Products Project"
author: "Eamon Corr"
date: "21 January 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Would you Survive the sinking of the Titanic?
Simply enter a few details and the Shiny App will predict your probability of survival with a 94% Accuracy.  

Experiment with different scenarios and see if your chances improve or deteriorate.

Click here for the Shiny App:
https://corralytics.shinyapps.io/titanic/

Follow the instructions below and Good Luck

## Instructions:
Open the Shiny App and enter the following details:

- **Sex:** *Male or Female*  
- **Ticket Class:** *1st, 2nd, 3rd*  
- **Embarkation:** *Cherbourg(France), Queenstown(Now Cobh, Ireland), S'hampton(England)*  
- **Age:** *For a child less than 1 enter 0.4*  
- **Parent/Child:** *Number of your parents and children that accompanied you on your journey*  
- **Family Size:** *Total number of family that accompanied you on the journey - include yourself in the total*

#### Additional guidelines are contained in the Shiny App input panel 

## The Titanic Training Dataset
The dataset used to build the prediction model looks like this:
```{r, comment = ""}
train <- read.csv("dataset/titanicShiny_train.csv")
test <- read.csv("dataset/titanicShiny_test.csv")
head(train,2)
```


Here's the calculated accuracy of the Model using a test dataset:
```{r, echo = TRUE}
model <- glm(formula=Survived ~. , family = binomial(link = "logit"), 
             data = train)
fitted.probs <- predict(model, newdata=test, type = "response")
fitted.pred <- ifelse(fitted.probs > 0.5, 1, 0)
accuracy <- mean(fitted.pred==test$Survived)

```

```{r}
accuracy
```


##Shiny App Documentation & Reproducibility

The Titanic Survival Predictor Shiny App is located here:
https://corralytics.shinyapps.io/titanic/

The full source code for this Shiny App can be found on my GitHub page: 
https://github.com/Eamoned/Titanic-Survival-Predictor-ShinyApp

All code for the initial titanic data exploration, data cleaning, feature engineering
& model building can be found here:  
https://github.com/Eamoned/Titanic-Survival-Predictor-ShinyApp/blob/master/titanicModelBuilding.rmd

Enjoy

